---
title: "Biodiversity data: rgbif data cleaning"
author: Scott Chamberlain
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

## rgbif data quality/data cleaning

```{r}
library('rgbif')
```

## Parsing output by issue

```{r}
(res <- occ_search(geometry='POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 50))
```

Can print whole table, or search for matches

```{r}
head(gbif_issues())
gbif_issues()[ gbif_issues()$code %in% c('cdround','cudc','gass84','txmathi'), ]
```

Or parse issues in various ways, e,g. remove data rows with certain issue classes

```{r}
library('magrittr')
res %>% occ_issues(-gass84, -mdatunl)
```

split issues into separate columns

```{r}
res %>% occ_issues(mutate = "split")
```

expand issues to more descriptive names

```{r}
res %>% occ_issues(mutate = "expand")
```

split and expand

```{r}
res %>% occ_issues(mutate = "split_expand")
```

split, expand, and remove an issue class

```{r}
res %>% occ_issues(-gass84, mutate = "split_expand")
```


## Other cleaning

* Contextual cleaning based on 
  * impossible lat/long values
  * expected habitat type
  * expected range of values
  * etc.
  

## Your turn

- 5 min

> Get some data from rgbif::occsearch, then use rgbif::occ_issues to 
parse data. 
