---
title: "Biodiversity data: rgbif downloads"
author: Scott Chamberlain
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

## rgbif data quality/data cleaning

```{r}
library('rgbif')
```

## Downloads

The user interface is slightly different from `occ_search`/`occ_data` and 
other functions.

Here, we query for records that have a `basisOfRecord` value of `LITERATURE`. 

```{r eval=FALSE}
x <- occ_download("basisOfRecord = LITERATURE")
```

That kicks off a request, and we get back basically an identifier with 
which we can check when the request is done, and fetch it. 

```{r eval=FALSE}
occ_download_meta(x)
```

Then when done, we can fetch and import in one line

```{r eval=FALSE}
occ_download_get(x) %>% occ_download_import()
```



## Your turn

- 5 min

> If you have GBIF credentials already, try to run a small download - 
to be sure it will be done very quickly, OR if you have a download you've 
already done on GBIF, you can use that ID to download that data with 
rgbif. 
