---
title: "Biodiversity data: spocc"
author: Scott Chamberlain
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

spocc - Make some maps!

## Load libraries

```{r}
library('spocc')
```

## spocc unifies access to biodiversity data across sources

```{r sources, message=FALSE, warning=FALSE}
out <- occ(query = 'Accipiter striatus', from = 'gbif')
out$gbif # GBIF data w/ metadata
out$ebird$data # empty
out$gbif$meta #  metadata, your query parameters, time the call executed, etc. 
out$gbif$data # just data
```

And you can squash together data from sources easily

```{r manysources, message=FALSE, warning=FALSE}
out <- occ(query = 'Accipiter striatus', from = c('gbif', 'ebird'))
occ2df(out)
```

## Your turn

- 5 min

> 1. Explore getting data from other sources for the same taxa to see 
how they compare. And combine output (hint: occ2df() function)
> 2. If you have time, explore other ways to query with spocc, including
possibly by spatial area, sp class spatial objects, taxize identifiers,
or other things.
