---
title: "Taxonomy: hierarchies"
author: Scott Chamberlain
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

## Hierarchies

Full hierarchies

```{r getdata}
library('taxize')
splist <- c("Helianthus annuus","Pinus contorta","Collomia grandiflora","Abies magnifica","Rosa californica","Datura wrightii","Mimulus bicolor","Nicotiana glauca","Madia sativa","Bartlettia scaposa")
class_list <- classification(splist, db = "gbif")
```

We can combine the hierarchy outputs

```{r manipulate}
(class_df <- cbind(class_list))
```

Now we can start seeing some relationships among species simply by their
shared taxonomic names

```{r index}
class_df[1:2, ]
```

## downstream

We can get downstream names, in this case all species from the genus *Apis*

```{r}
downstream("Apis", db = "col", downto = "Species")
```

## upstream

We can also upstream names, in this case all genera up genus *Apis*

```{r eval=FALSE}
upstream('Poa annua', db = 'col', upto = 'family')
#> $`Poa annua`
#>                        childtaxa_id   childtaxa_name childtaxa_rank
#> 1  3834e28670fe3c8f76fea24fb496a15f    Anarthriaceae         family
#> 2  0cead872c131282b96da3b3ca3a6fbf8     Bromeliaceae         family
#> 3  4ec08d1396898b750fe1727545ac273b Centrolepidaceae         family
#> 4  9c6d72055c3933fb772878569d1e2744       Cyperaceae         family
#> 5  b11f7be500656ec5ea365a1c7a07a35f  Ecdeiocoleaceae         family
#> 6  0d9601b89e891871a79b72d28b00f1c5    Eriocaulaceae         family
#> 7  dcb2312860f7923e219fc4f133576021  Flagellariaceae         family
#> 8  4e33ddf845e917d310361e74cfb39370   Joinvilleaceae         family
#> 9  d2bce625cc55cf37f6cbc1143ad48047        Juncaceae         family
#> 10 2c3053c3692654b0c949ccf466b1c12e       Mayacaceae         family
#> 11 d01bb56e344e74c4cd8279d28612897e          Poaceae         family
#> 12 529357780bc2958682ac01662c0d6bc3      Rapateaceae         family
#> 13 2d03ecec028bdd7142736243105e905c     Restionaceae         family
#> 14 8f16a0a86bc6377af6cbfba51fe89c38      Thurniaceae         family
#> 15 6d55aefffa2cb553f22f0758507be694        Typhaceae         family
#> 16 a31c53aa2ddbf8f86f17d5b89ffe2406       Xyridaceae         family
```

## Your turn

- 5 min

> For 10 names, search by name or get taxonomic IDs using our previous tools, 
and get full hierarchy, and bind them together both row-wise and column-wise.
