---
title: "Taxonomy: identifiers"
author: Scott Chamberlain
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

# Taxonomic identifiers

## Search by taxonomic name

```{r}
library('taxize')
(out <- get_uid("Poa annua"))
```

## Sometimes you are presented with a prompt

```{r eval=FALSE}
get_colid("Alces alces")
#>                       name         rank                            colid   name_status  kingdom   family
#> 1              Alces alces      species fc7d998315ba4a3e50cfab10954bdc45 accepted name Animalia Cervidae
#> 2        Alces alces alces infraspecies f2d7d57ae493db1767049fc11bb999cf accepted name Animalia Cervidae
#> 3   Alces alces caucasicus infraspecies ffdb4bc3627d6293e48b07d67802388d accepted name Animalia Cervidae
#> 4    Alces alces americana infraspecies 505aab1a2067afac631af51b575d115e       synonym     <NA>     <NA>
#> 5    Alces alces andersoni infraspecies 505aab1a2067afac631af51b575d115e       synonym     <NA>     <NA>
#> 6   Alces alces cameloides infraspecies c8557a9c74530ec84395efaaec4f6bfd       synonym     <NA>     <NA>
#> 7        Alces alces gigas infraspecies 505aab1a2067afac631af51b575d115e       synonym     <NA>     <NA>
#> 8 Alces alces pfizenmayeri infraspecies 505aab1a2067afac631af51b575d115e       synonym     <NA>     <NA>
#> 9      Alces alces shirasi infraspecies 505aab1a2067afac631af51b575d115e       synonym     <NA>     <NA>
```

## Underscore method: get all data 

```{r}
get_colid_("Alces alces")
```

Then choose which name yourself, ideally programatically.


## Coerce ids to ID classes

```{r}
as.gbifid(2704179)
```

```{r}
as.gbifid(2704179, check = FALSE)
```

Then many functions become available that act on these ID classes

## Many names - the ids class

In this case get many identifiers for many names, then pass in identifiers to get many classifications

```{r}
mynames <- c("Helianthus annuus","Pinus contorta","Collomia grandiflora")
out <- get_ids(mynames, db = c('ncbi','itis','col','tropicos'))
out$tropicos
class(out)
class(out$col)
classification(out$col)
```


## Your turn

- 5 min

> Search 3 different data sources via get* functions, each with a few
taxonomic names you know off the top of your head. If you can't think of
any see names_list() function.



# Use identifiers with other functions

Taxonomic hierarchy

```{r}
classification(9606, db = 'ncbi')
```


Get synonyms

```{r}
synonyms('Pomatomus saltatrix', db = 'worms')
```


Get children

```{r}
children("Salmo", db = 'worms')
```


# Your turn

- 5 min

> Take some taxonomic IDs or get new ones, and inputs those IDs to other
taxize functions. Try to use one above you just saw and another one 
in the documentation. 
